Comparison of Global Solution Methods to a Zero Lower Bound Model

### 2019-04-25 Presentation/Presentation.tex
Add in clicking bullets

\begin{itemize}\setlength{\itemsep}{7pt}
  \item  As a result of the global financial crisis of 2007-9 and the subsequent recession, central banks lowered their policy rate to its zero lower bound (ZLB)
  \item  The ZLB now holds for a significant portion of historic data for the US, Japan, and the Euro Area
  \item  The ZLB introduces a kink in the central bank's policy rule and calls into question linear estimation methods
  \item  Responses in the literature to this nonlinearity include:
  %\item This paper:
  \begin{enumerate}\setlength{\itemsep}{4pt}
	\item  Failing to incorporate ZLB period data
	\item  Estimating linear models on the entire data set
	\item  Estimating a piecewise linear version of the nonlinear model (e.g., Guerrieri and Iacoviello, 2017)
	\item  Estimating fully nonlinear models that treats the ZLB as an occasionally binding constraint (e.g., Gust et al., 2017; Plante et al., 2018; Richter and Throckmorton, 2016).
  \end{enumerate}
\end{itemize}
\end{frame}

@@

\begin{frame}\frametitle{Nonlinear Solution Methods}
\begin{itemize}\setlength{\itemsep}{10pt}
	\item  As it is important to incorporate all historical data and yield acurate parameter values and predictions, the path forward seems to be in fully nonlinear models
	\item Solving nonlinear models with projection methods (e.g. Aruoba et al., 2018; Ferna\'ndez-Villaverde et al., 2015).
	\item  Researchers have used different methods for solving fully nonlinear models:
	\begin{enumerate} \setlength{\itemsep}{4pt}
		\item  Policy function iteration with linear interpolation (e.g., Plante et al., 2018; Richter and Throckmorton, 2016).
		\item  Alternate approximating functions: regime-indexed policy functions (Gust et al., 2017), piecewise smooth policy functions (Aruoba et al. 2018)
		\item  Alternate grid construction: Smolyak method with Chebyshev polynomials (e.g., Gust et al., 2017; Ferna\'ndez-Villaverde et al. 2015; Aruoba et al., 2018)
	\end{enumerate}
\end{itemize}
\end{frame}

%=========================================================================
\begin{frame}\frametitle{Two Solution Methods}
\begin{table}\scriptsize
    \renewcommand*{\arraystretch}{1.4}
    \setlength{\tabcolsep}{6pt}      
  \begin{tabularx}{\textwidth}{l | X | X}
    \hline
    & \textbf{Atkinson et al.\ (2019)} & \textbf{Gust et al.\ (2017)} \\ \hline
    	Discretization Method & Evenly-spaced grid points & Smolyak Method\\
	Policy Function Evaluation & Linear Interpolation & Chebyshev Polynomials \\
	Integration Method & Rouwenhorst & Gauss-Hermite quadrature  \\ \hline
\end{tabularx}
\end{table}\normalsize
\begin{itemize}\setlength{\itemsep}{10pt}
\item  Evenly-spaced grid points work well with linear interpolation
\item  The Smolyak method is optimal for Chebyshev polynomials
\item  The Rouwenhorst (1995) method improves approximation on exogenous dimensions when the driving processes are autoregressive
\end{itemize}

\end{frame}
%=========================================================================
\begin{frame}\frametitle{Atkinson et al.\ (2019)}
\begin{itemize}\setlength{\itemsep}{10pt}
\item  Time iteration with linear interpolation
\item Directly approximates policy functions
\item  State space features evenly-spaced nodes
\item  Exogenous variables are approximated with the Markov chain from Rouwenhorst (1995)
  \begin{itemize}\setlength{\itemsep}{4pt}
\item Only requires interpolation along the dimensions of the endogenous state variables
\end{itemize}
\end{itemize}
\end{frame}
%=========================================================================
\begin{frame}\frametitle{Gust et al.\ (2017)}
\begin{itemize}\setlength{\itemsep}{6pt}
\item  Approximates policy functions using an antistropic Smolyak method with Chebyshev polynomials
\item  Approximates exogenous state variables using Gauss-Hermite quadrature%why if we know Rouwenhorst is better?
\item  Instead of directly computing the policy functions, they estimate functions at and away from the ZLB, building on Christiano and Fisher (2000)
  \begin{itemize}\setlength{\itemsep}{4pt}
\item  Policy functions feature a kink or non-differentiability at the ZLB and regime-indexing the policy functions will yield smoother functions
\item  Low-order Chebyshev polynomials perform better on smoother functions and are inexpensive to compute
\end{itemize}
\end{itemize}
\end{frame}
%=========================================================================
\begin{frame}\frametitle{This Project}
\begin{itemize}\setlength{\itemsep}{8pt}
\item  In this project, we consider how splitting up the policy functions conditional on the ZLB impacts the speed and accuracy of the solution of a nonlinear model
\item  We use policy function iteration on a fixed point with linear interpolation
\item  Atkinson et al.\ (2019) provides the motivation for a single policy function;  Gust et al.\ (2017) provides the motivation for a regime-indexed policy function
\item  Solution algorithm based on Atkinson et al.\ (2019) denoted henceforth as ART; solution algorithm based on Gust et al.\ (2017) denoted henceforth as GHLS
\end{itemize}
\end{frame}
%=========================================================================
\begin{frame}\frametitle{Our Solution Algorithm} %NEED TO PUT SOME ABOUT GHLS?
\begin{itemize}\setlength{\itemsep}{8pt}
\item   Construct the state space with evenly-spaced nodes and approximate exogenous state variables with Rouwenhorst (1995)
\item  Obtain initial conjectures for a set of policy functions from the log-linear solution
\item  Using a fixed point iteration scheme, linearly interpolate and numerically integrate the policy functions each step
\begin{itemize}
\item  For ART portion, we update a single policy function, and for the GHLS portion we update regime-indexed policy functions
\end{itemize}
\item  The algorithm converges once the maximum distance between successive guesses of policy functions falls below a convergence criterion

@@

\item  The interest rate enters directly in the consumption Euler equation

@@

\item  At the ZLB, the presence of the interest rate creates a nonlinearity in the consumption Euler equation 
\item  $c_t$ depends directly on the interest rate

@@

\item The households choose $\{c_t, n_t, b_t, x_t, k_t\}_{t=0}^\infty$ to maximize expected lifetime utility given by

@@

\item The nominal bond, $b$, is subject to a risk premium, $s$, that follows

@@

\item Households also face an investment adjustment cost, so the law of motion for capital is given by

@@

\item The FOCs to each household's constrained optimization problem are

@@

\item The production sector consists of a continuum of monopolistically competitive intermediate goods firms and a final goods firm 
\item Technology is $z_t = g_tz_{t-1}$, which is common across firms
\item Deviations from the steady-state growth rate, $\bar{g}$, follow

@@

\item In symmetric equilibrium, the optimality conditions reduce to

@@

\item

@@

\item A more negative net notional rate indicates that the central bank is more constrained

@@

%=========================================================================
\begin{frame}\frametitle{Competitive Equilibrium}\setlength{\itemindent}{-1.2em}
\begin{itemize} \setlength{\itemsep}{8pt}
\item The aggregate resource constraint and real GDP definition are given by:
\begin{gather*}
  c_t + x_t = y_t^{gdp}\\
  y_t^{gdp} = [1 - \varphi(\pi_t^{gap}-1)^2/2]y_t
\end{gather*}
\item The model does not have a steady-state due to the unit root in technology, $z_t$. Therefore, we define the variables with a trend in terms of technology (i.e., $\tilde{x}_t \equiv x_t/z_t$)
\end{itemize}
\end{frame}
%=========================================================================
\begin{frame}\frametitle{Competitive Equilibrium}
\begin{itemize} \setlength{\itemsep}{10pt}
\item A competitive equilibrium consists of sequences of 
\begin{enumerate} \setlength{\itemsep}{5pt}
\item quantities, $\{\tilde{c}_t, \tilde{y}_t, \tilde{y}_t^{gdp}, x^g_t, y^g_t, n_t, \tilde{k}_t, \tilde{x}_t\}_{t=0}^\infty$ \item prices, $\{\tilde{w}_t, i_t, i^n_t, \pi_t, \tilde{\lambda}_t, q_t, r^k_t, mc_t\}_{t=0}^\infty$
\item exogenous variables, $\{s_t,g_t\}_{t=0}^\infty$
\end{enumerate}
\item that satisfy the detrended equilibrium system, given 
\begin{enumerate}\setlength{\itemsep}{5pt}
\item the initial conditions, $\{\tilde{c}_{-1}, i^n_{-1}, \tilde{k}_{-1}, \tilde{x}_{-1}, \tilde{w}_{-1}, s_0, g_0, \varepsilon_{i,0}\}$
\item three sequences of shocks, $\{\varepsilon_{g,t}, \varepsilon_{s,t}, \varepsilon_{i,t}\}_{t=1}^\infty$
\end{enumerate}
\end{itemize}
\end{frame}
%===============================================================================================================
\begin{frame}\frametitle{Parameter Values}
\begin{table}[H]
\centering
\captionsetup{justification=centering}
  \scriptsize
    \setlength{\tabcolsep}{3pt}      
  \begin{tabular}{l c c |l  c c}
    \hline
  Subjective Discount Factor & $\beta$ & $0.9949$ & Rotemberg Price Adjustment Cost & $\varphi$ & $100$ \\
   Frisch Labor Supply Elasticity & $1/\eta$ & $3$ & Inflation Gap Response & $\phi_\pi$ & $2.0$ \\
  Price Elasticity of Substitution & $\theta$ & $6$ & Output Growth Gap Response & $\phi_y$ & $0.5$ \\
  Steady-State Labor Hours & $\bar{n}$ & $1/3$ & Habit Persistence & $h$ & $0.80$ \\
  Steady-State Risk Premium & $\bar{s}$ & $1.0058$ & Risk Premium Persistence & $\rho_s$ & $0.80$ \\
  Steady-State Growth Rate & $\bar{g}$ & $1.0034$ & Notional Rate Persistence & $\rho_i$ & $0.80$ \\
  Steady-State Inflation Rate & $\bar{\pi}$ & $1.0053$ & Technology Growth Shock SD & $\sigma_g$ & $0.005$ \\
  Capital Share of Income & $\alpha$ & $0.35$ & Risk Premium Shock SD & $\sigma_s$ & $0.0085$ \\
  Capital Depreciation Rate & $\delta$ & $0.025$ & Notional Interest Rate Shock SD & $\sigma_i$ & $0.002$ \\
  Investment Adjustment Cost & $\nu$ & $4$ & & &\\
  \hline
  \end{tabular}
\normalfont
\end{table}
\begin{itemize}\setlength{\itemsep}{8pt}
\item Parameters are from Atkinson et al.\ (2019), and are chosen to be characteristic of U.S. data
\item For the model without capital, $\rho_s = 0.006$ %To allow the GHLS algorithm to converge
\end{itemize}
\end{frame}
%===================================================================================
\begin{frame}\frametitle{Discretized State Space (No Capital)}

\begin{itemize}\setlength{\itemsep}{10pt}
  \item State variables: $g_t$, $s_t$, $mp_t$, $i^n_{t-1}$
  \item Number of grid points: $N_g$, $N_{s}$, $N_{mp}$, $N_{i^n}$
  \item Grid boundaries: 
\begin{center}
$[g_{\min},g_{\max}]$, $[s_{\min},s_{\max}]$, $[mp_{\min},mp_{\max}]$, $[i^n_{\min},i^n_{\max}]$
\end{center}
  \item Create evenly spaced grids:
  \begin{center}
    $x_{grid} = \texttt{linspace}(x_{\min},x_{\max},N_x), \quad x\in\{g,s,mp,i^n\}$
  \end{center}
  \item State space contains $N=N_g\times N_s\times N_{mp}\times N_{i^n}$ nodes
  \item Create an array for each state variable, where every position is a unique permutation of the state space:
  \begin{center}
    $[g_{gr},s_{gr},mp_{gr},i^n_{gr}] = \texttt{ndgrid}(g_{grid},s_{grid},mp_{grid},i^n_{grid})$
  \end{center}
\end{itemize}

@@

\begin{itemize}\setlength{\itemsep}{12pt}
  \item True RE solution only exists in special cases %(e.g., $\delta=1$)
  \item Goal: Find an approximating function that maps the state space to the optimal decision rule for consumption:
  \begin{gather*}
    \underbrace{c(g,s,mp,i^n)}_{\textrm{True RE Solution}} \approx \underbrace{{\cal{P}}_c(g,s,mp,i^n)}_{\textrm{Approximating Function}}
  \end{gather*}
  \item Basic elements of the algorithm:
  \begin{enumerate}\setlength{\itemsep}{6pt}
    \item Interpolation: Linear, Chebyshev polynomials %Leasst Squares
    \item Integration: Rouwenhorst, Gauss-Hermite %Trapezoid, 
    \item Iteration: Time, Fixed-point 
  \end{enumerate}
\end{itemize}

@@

\begin{itemize}\setlength{\itemsep}{8pt}
  \item Piecewise Linear Interpolation: 4 state variables $(g,s,mp,i^n)$
  \item Goal: Find the policy function value ${\cal{P}}_c(g',s',mp',{i^n}')$
  \item We have policy function values on nearest 24 nodes
  \begin{gather*}
   [{\cal{P}}_c(g_i,s_j,mp_k,i^n_l), {\cal{P}}_c(g_i,s_j,mp_k,i^n_{l+1}), {\cal{P}}_c(g_i,s_j,mp_{k+1},i^n_l), \\ \dots, {\cal{P}}_c(g_{i+1},s_{j+1},mp_{k+1},i^n_{l+1})]
  \end{gather*}
  once we determine the grid indices, $i,j,k,l$
  \item Locate the grid point left of $x'$, $x \in \{g,s,mp,i^n\}$ %nearest point to the left
   \begin{gather*}
    \textrm{step} = x_2 - x_1, \qquad \textrm{dist} = x' - x_1\\
    \textrm{loc} = \min(N_x-1,\max(1,\operatorname{floor}(\textrm{dist}/\textrm{step}) + 1))
    \end{gather*}
\end{itemize}

@@

\begin{itemize}\setlength{\itemsep}{10pt}
  \item A general class of polynomials can be written as:
  \begin{gather*}
  {\cal{P}}(x;\eta) = \sum_{i=0}^n\eta_i\varphi_i(x).
  \end{gather*}
  \item Linear interpolation is a special case of this general class (i.e., $n=1$, $\varphi_i(x)=x^i$, and weights chosen appropriately)
  \item Can also use bases consisting of orthogonal polynomials (e.g., Chebyshev Polynomials)
\end{itemize}

@@

\begin{enumerate}\setlength{\itemsep}{10pt}
  \item Calculate $z(s')$ for each realization of the state
  \item Find the updated policy function, $n(s')$, for each state
  \item Numerical Integration:
  \begin{itemize}\setlength{\itemsep}{5pt}
    \item First integrate across the continuous random variable, $z$, conditional on the future realizations of the discrete stochastic variable, $s'$.%, to obtain:
        %\begin{gather*}
         % E[\Phi(z')|s'=1], \quad E[\Phi(z')|s'=2]
        %\end{gather*}
    \item Then weight the conditional expectations by their corresponding likelihood. %The conditional expectations are:
  %  \begin{gather*}
    %  E\left[\Phi(z')|s = i\right] = p_{i1}E[\Phi(z')|s'=1] + p_{i2}E[\Phi(z')|s'=2]
   % \end{gather*}
  \end{itemize}
\end{enumerate}

@@

\begin{itemize}\setlength{\itemsep}{8pt}
  \item Used to approximate an exogenous $AR(1)$ process
  \item Kopecky and Suen (2010) show the Rouwenhorst method outperforms other approximations of an $AR(1)$ process
  \item The approximation is a Markov switching process like the time-varying intercept example, but with $n$ states
  \item The method determines the bounds of the exogenous state variables, the nodes, and the transition probabilities
  \item Let $z\sim AR(1)$ with persistence $\rho$, mean $\mu_z$, and variance %Idea for approximation of an AR(1) process
  \begin{gather*}
    \sigma_z^2 = \sigma_\varepsilon^2/(1-\rho^2).
  \end{gather*}
  \item The $n$ states for the discretized process $z$ are evenly spaced on $[\mu_z-\sigma_z\sqrt{n-1},\mu_z+\sigma_z\sqrt{n-1}]$
\end{itemize}

@@

\begin{itemize}\setlength{\itemsep}{14pt}
  \item Any calculations that are not dependent on the results of other calculations can be performed in parallel (e.g., solving for policy values at each node in the state space)
  \item We replace \texttt{for} loops with \texttt{parfor} loops where applicable. This tells MATLAB to distribute each step in the loop across the specified number of processors
\end{itemize}

@@

%===============================================================================================================
\begin{frame}\frametitle{Conclusion}
\begin{itemize}\setlength{\itemsep}{10pt}
\item Key takeaways:
\begin{itemize}\setlength{\itemsep}{4pt}
\item In the model without capital, regime-indexed policy functions were smoother and solution algorithm was faster
\item In model with capital,  the regime-indexed policy functions were more nonlinear and solution algorithm was slower
\end{itemize}
\item Extensions:
\begin{itemize}\setlength{\itemsep}{4pt}
\item Solve the GHLS solution method using Smolyak discretization methods and Chebyshev polynomials 
\item This would be expected to speed up the solution algorithm, but may lead to a less accurate solution
\end{itemize}
\end{itemize}
\end{frame}

>>>>

\begin{itemize}\setlength{\itemsep}{7pt}
  \item <1-|handout:1>As a result of the global financial crisis of 2007-9 and the subsequent recession, central banks lowered their policy rate to its zero lower bound (ZLB)
  \item  <2-|handout:1>The ZLB now holds for a significant portion of historic data for the US, Japan, and the Euro Area
  \item  <3-|handout:1>The ZLB introduces a kink in the central bank's policy rule and calls into question linear estimation methods
  \item <4-|handout:1> Responses in the literature to this nonlinearity include:
  %\item This paper:
  \begin{enumerate}\setlength{\itemsep}{4pt}
	\item  Failing to incorporate ZLB period data
	\item  Estimating linear models on the entire data set
	\item  Estimating a piecewise linear version of the nonlinear model (e.g., Guerrieri and Iacoviello, 2017)
	\item  Estimating fully nonlinear models that treats the ZLB as an occasionally binding constraint (e.g., Gust et al., 2017; Plante et al., 2018; Richter and Throckmorton, 2016).
  \end{enumerate}
\end{itemize}
\end{frame}

@@

\begin{frame}\frametitle{Nonlinear Solution Methods}
\begin{itemize}\setlength{\itemsep}{10pt}
	\item  <1-|handout:1>As it is important to incorporate all historical data and yield acurate parameter values and predictions, the path forward seems to be in fully nonlinear models
	\item <2-|handout:1>Solving nonlinear models with projection methods (e.g. Aruoba et al., 2018; Ferna\'ndez-Villaverde et al., 2015).
	\item  <3-|handout:1>Researchers have used different methods for solving fully nonlinear models:
	\begin{enumerate} \setlength{\itemsep}{4pt}
		\item  Policy function iteration with linear interpolation (e.g., Plante et al., 2018; Richter and Throckmorton, 2016).
		\item  Alternate approximating functions: regime-indexed policy functions (Gust et al., 2017), piecewise smooth policy functions (Aruoba et al. 2018)
		\item  Alternate grid construction: Smolyak method with Chebyshev polynomials (e.g., Gust et al., 2017; Ferna\'ndez-Villaverde et al. 2015; Aruoba et al., 2018)
	\end{enumerate}
\end{itemize}
\end{frame}

%=========================================================================
\begin{frame}\frametitle{Two Solution Methods}
\begin{table}\scriptsize
    \renewcommand*{\arraystretch}{1.4}
    \setlength{\tabcolsep}{6pt}      
  \begin{tabularx}{\textwidth}{l | X | X}
    \hline
    & \textbf{Atkinson et al.\ (2019)} & \textbf{Gust et al.\ (2017)} \\ \hline
    	Discretization Method & Evenly-spaced grid points & Smolyak Method\\
	Policy Function Evaluation & Linear Interpolation & Chebyshev Polynomials \\
	Integration Method & Rouwenhorst & Gauss-Hermite quadrature  \\ \hline
\end{tabularx}
\end{table}\normalsize
\begin{itemize}\setlength{\itemsep}{10pt}
\item  <2-|handout:1>Evenly-spaced grid points work well with linear interpolation
\item  <3-|handout:1>The Smolyak method is optimal for Chebyshev polynomials
\item  <4-|handout:1>The Rouwenhorst (1995) method improves approximation on exogenous dimensions when the driving processes are autoregressive
\end{itemize}

\end{frame}
%=========================================================================
\begin{frame}\frametitle{Atkinson et al.\ (2019)}
\begin{itemize}\setlength{\itemsep}{10pt}
\item <1-|handout:1> Time iteration with linear interpolation
\item <2-|handout:1>Directly approximates policy functions
\item <3-|handout:1> State space features evenly-spaced nodes
\item<4-|handout:1>  Exogenous variables are approximated with the Markov chain from Rouwenhorst (1995)
  \begin{itemize}\setlength{\itemsep}{4pt}
\item Only requires interpolation along the dimensions of the endogenous state variables
\end{itemize}
\end{itemize}
\end{frame}
%=========================================================================
\begin{frame}\frametitle{Gust et al.\ (2017)}
\begin{itemize}\setlength{\itemsep}{6pt}
\item  <1-|handout:1>Approximates policy functions using an antistropic Smolyak method with Chebyshev polynomials
\item  <2-|handout:1>Approximates exogenous state variables using Gauss-Hermite quadrature%why if we know Rouwenhorst is better?
\item  <3-|handout:1>Instead of directly computing the policy functions, they estimate functions at and away from the ZLB, building on Christiano and Fisher (2000)
  \begin{itemize}\setlength{\itemsep}{4pt}
\item  Policy functions feature a kink or non-differentiability at the ZLB and regime-indexing the policy functions will yield smoother functions
\item Low-order Chebyshev polynomials perform better on smoother functions and are inexpensive to compute
\end{itemize}
\end{itemize}
\end{frame}
%=========================================================================
\begin{frame}\frametitle{This Project}
\begin{itemize}\setlength{\itemsep}{8pt}
\item  <1-|handout:1>In this project, we consider how splitting up the policy functions conditional on the ZLB impacts the speed and accuracy of the solution of a nonlinear model
\item  <2-|handout:1>We use policy function iteration on a fixed point with linear interpolation
\item  <3-|handout:1>Atkinson et al.\ (2019) provides the motivation for a single policy function;  Gust et al.\ (2017) provides the motivation for a regime-indexed policy function
\item <4-|handout:1> Solution algorithm based on Atkinson et al.\ (2019) denoted henceforth as ART; solution algorithm based on Gust et al.\ (2017) denoted henceforth as GHLS
\end{itemize}
\end{frame}
%=========================================================================
\begin{frame}\frametitle{Our Solution Algorithm} %NEED TO PUT SOME ABOUT GHLS?
\begin{itemize}\setlength{\itemsep}{8pt}
\item   <1-|handout:1>Construct the state space with evenly-spaced nodes and approximate exogenous state variables with Rouwenhorst (1995)
\item  <2-|handout:1>Obtain initial conjectures for a set of policy functions from the log-linear solution
\item  <3-|handout:1>Using a fixed point iteration scheme, linearly interpolate and numerically integrate the policy functions each step
\begin{itemize}
\item For ART portion, we update a single policy function, and for the GHLS portion we update regime-indexed policy functions
\end{itemize}
\item <5-|handout:1> The algorithm converges once the maximum distance between successive guesses of policy functions falls below a convergence criterion

@@

\item  <1-|handout:1>The interest rate enters directly in the consumption Euler equation

@@

\item <2-|handout:1> At the ZLB, the presence of the interest rate creates a nonlinearity in the consumption Euler equation 
\item <3-|handout:1> $c_t$ depends directly on the interest rate

@@

\item <1-|handout:1>The households choose $\{c_t, n_t, b_t, x_t, k_t\}_{t=0}^\infty$ to maximize expected lifetime utility given by

@@

\item <2-|handout:1>The nominal bond, $b$, is subject to a risk premium, $s$, that follows

@@

\item <1-|handout:1>Households also face an investment adjustment cost, so the law of motion for capital is given by

@@

\item <2-|handout:1>The FOCs to each household's constrained optimization problem are

@@

\item <1-|handout:1>The production sector consists of a continuum of monopolistically competitive intermediate goods firms and a final goods firm 
\item <2-|handout:1>Technology is $z_t = g_tz_{t-1}$, which is common across firms
\item <3-|handout:1>Deviations from the steady-state growth rate, $\bar{g}$, follow

@@

\item <4-|handout:1>In symmetric equilibrium, the optimality conditions reduce to

@@

\item <1-|handout:1>\ 

@@

\item <2-|handout:1>A more negative net notional rate indicates that the central bank is more constrained

@@

%=========================================================================
\begin{frame}\frametitle{Competitive Equilibrium}\setlength{\itemindent}{-1.2em}
\begin{itemize} \setlength{\itemsep}{8pt}
\item <1-|handout:1>The aggregate resource constraint and real GDP definition are given by:
\begin{gather*}
  c_t + x_t = y_t^{gdp}\\
  y_t^{gdp} = [1 - \varphi(\pi_t^{gap}-1)^2/2]y_t
\end{gather*}
\item <2-|handout:1>The model does not have a steady-state due to the unit root in technology, $z_t$. Therefore, we define the variables with a trend in terms of technology (i.e., $\tilde{x}_t \equiv x_t/z_t$)
\end{itemize}
\end{frame}
%=========================================================================
\begin{frame}\frametitle{Competitive Equilibrium}
\begin{itemize} \setlength{\itemsep}{10pt}
\item <1-|handout:1>A competitive equilibrium consists of sequences of 
\begin{enumerate} \setlength{\itemsep}{5pt}
\item quantities, $\{\tilde{c}_t, \tilde{y}_t, \tilde{y}_t^{gdp}, x^g_t, y^g_t, n_t, \tilde{k}_t, \tilde{x}_t\}_{t=0}^\infty$ \item prices, $\{\tilde{w}_t, i_t, i^n_t, \pi_t, \tilde{\lambda}_t, q_t, r^k_t, mc_t\}_{t=0}^\infty$
\item exogenous variables, $\{s_t,g_t\}_{t=0}^\infty$
\end{enumerate}
\item <2-|handout:1>that satisfy the detrended equilibrium system, given 
\begin{enumerate}\setlength{\itemsep}{5pt}
\item the initial conditions, $\{\tilde{c}_{-1}, i^n_{-1}, \tilde{k}_{-1}, \tilde{x}_{-1}, \tilde{w}_{-1}, s_0, g_0, \varepsilon_{i,0}\}$
\item three sequences of shocks, $\{\varepsilon_{g,t}, \varepsilon_{s,t}, \varepsilon_{i,t}\}_{t=1}^\infty$
\end{enumerate}
\end{itemize}
\end{frame}
%===============================================================================================================
\begin{frame}\frametitle{Parameter Values}
\begin{table}[H]
\centering
\captionsetup{justification=centering}
  \scriptsize
    \setlength{\tabcolsep}{3pt}      
  \begin{tabular}{l c c |l  c c}
    \hline
  Subjective Discount Factor & $\beta$ & $0.9949$ & Rotemberg Price Adjustment Cost & $\varphi$ & $100$ \\
   Frisch Labor Supply Elasticity & $1/\eta$ & $3$ & Inflation Gap Response & $\phi_\pi$ & $2.0$ \\
  Price Elasticity of Substitution & $\theta$ & $6$ & Output Growth Gap Response & $\phi_y$ & $0.5$ \\
  Steady-State Labor Hours & $\bar{n}$ & $1/3$ & Habit Persistence & $h$ & $0.80$ \\
  Steady-State Risk Premium & $\bar{s}$ & $1.0058$ & Risk Premium Persistence & $\rho_s$ & $0.80$ \\
  Steady-State Growth Rate & $\bar{g}$ & $1.0034$ & Notional Rate Persistence & $\rho_i$ & $0.80$ \\
  Steady-State Inflation Rate & $\bar{\pi}$ & $1.0053$ & Technology Growth Shock SD & $\sigma_g$ & $0.005$ \\
  Capital Share of Income & $\alpha$ & $0.35$ & Risk Premium Shock SD & $\sigma_s$ & $0.0085$ \\
  Capital Depreciation Rate & $\delta$ & $0.025$ & Notional Interest Rate Shock SD & $\sigma_i$ & $0.002$ \\
  Investment Adjustment Cost & $\nu$ & $4$ & & &\\
  \hline
  \end{tabular}
\normalfont
\end{table}
\begin{itemize}\setlength{\itemsep}{8pt}
\item <2-|handout:1>Parameters are from Atkinson et al.\ (2019), and are chosen to be characteristic of U.S. data
\item <3-|handout:1>For the model without capital, $\rho_s = 0.006$ %To allow the GHLS algorithm to converge
\end{itemize}
\end{frame}
%===================================================================================
\begin{frame}\frametitle{Discretized State Space (No Capital)}

\begin{itemize}\setlength{\itemsep}{10pt}
  \item <1-|handout:1>State variables: $g_t$, $s_t$, $mp_t$, $i^n_{t-1}$
  \item <2-|handout:1>Number of grid points: $N_g$, $N_{s}$, $N_{mp}$, $N_{i^n}$
  \item <3-|handout:1>Grid boundaries: 
\begin{center}
$[g_{\min},g_{\max}]$, $[s_{\min},s_{\max}]$, $[mp_{\min},mp_{\max}]$, $[i^n_{\min},i^n_{\max}]$
\end{center}
  \item <4-|handout:1>Create evenly spaced grids:
  \begin{center}
    $x_{grid} = \texttt{linspace}(x_{\min},x_{\max},N_x), \quad x\in\{g,s,mp,i^n\}$
  \end{center}
  \item <5-|handout:1>State space contains $N=N_g\times N_s\times N_{mp}\times N_{i^n}$ nodes
  \item <6-|handout:1>Create an array for each state variable, where every position is a unique permutation of the state space:
  \begin{center}
    $[g_{gr},s_{gr},mp_{gr},i^n_{gr}] = \texttt{ndgrid}(g_{grid},s_{grid},mp_{grid},i^n_{grid})$
  \end{center}
\end{itemize}

@@

\begin{itemize}\setlength{\itemsep}{12pt}
  \item <1-|handout:1>True RE solution only exists in special cases %(e.g., $\delta=1$)
  \item <2-|handout:1>Goal: Find an approximating function that maps the state space to the optimal decision rule for consumption:
  \begin{gather*}
    \underbrace{c(g,s,mp,i^n)}_{\textrm{True RE Solution}} \approx \underbrace{{\cal{P}}_c(g,s,mp,i^n)}_{\textrm{Approximating Function}}
  \end{gather*}
  \item <3-|handout:1>Basic elements of the algorithm:
  \begin{enumerate}\setlength{\itemsep}{6pt}
    \item Interpolation: Linear, Chebyshev polynomials %Leasst Squares
    \item Integration: Rouwenhorst, Gauss-Hermite %Trapezoid, 
    \item Iteration: Time, Fixed-point 
  \end{enumerate}
\end{itemize}

@@

\begin{itemize}\setlength{\itemsep}{8pt}
  \item <1-|handout:1>Piecewise Linear Interpolation: 4 state variables $(g,s,mp,i^n)$
  \item <2-|handout:1>Goal: Find the policy function value ${\cal{P}}_c(g',s',mp',{i^n}')$
  \item <3-|handout:1>We have policy function values on nearest 24 nodes
  \begin{gather*}
   [{\cal{P}}_c(g_i,s_j,mp_k,i^n_l), {\cal{P}}_c(g_i,s_j,mp_k,i^n_{l+1}), {\cal{P}}_c(g_i,s_j,mp_{k+1},i^n_l), \\ \dots, {\cal{P}}_c(g_{i+1},s_{j+1},mp_{k+1},i^n_{l+1})]
  \end{gather*}
  once we determine the grid indices, $i,j,k,l$
\item   <4-|handout:1>Locate the grid point left of $x'$, $x \in \{g,s,mp,i^n\}$ %nearest point to the left
   \begin{gather*}
    \textrm{step} = x_2 - x_1, \qquad \textrm{dist} = x' - x_1\\
    \textrm{loc} = \min(N_x-1,\max(1,\operatorname{floor}(\textrm{dist}/\textrm{step}) + 1))
    \end{gather*}
\end{itemize}

@@

\begin{itemize}\setlength{\itemsep}{10pt}
  \item <1-|handout:1>A general class of polynomials can be written as:
  \begin{gather*}
  {\cal{P}}(x;\eta) = \sum_{i=0}^n\eta_i\varphi_i(x).
  \end{gather*}
  \item <2-|handout:1>Linear interpolation is a special case of this general class (i.e., $n=1$, $\varphi_i(x)=x^i$, and weights chosen appropriately)
  \item <3-|handout:1>Can also use bases consisting of orthogonal polynomials (e.g., Chebyshev Polynomials)
\end{itemize}

@@

\begin{enumerate}\setlength{\itemsep}{10pt}
  \item <1-|handout:1>Calculate $z(s')$ for each realization of the state
  \item <2-|handout:1>Find the updated policy function, $n(s')$, for each state
  \item <3-|handout:1>Numerical Integration:
  \begin{itemize}\setlength{\itemsep}{5pt}
    \item First integrate across the continuous random variable, $z$, conditional on the future realizations of the discrete stochastic variable, $s'$.%, to obtain:
        %\begin{gather*}
         % E[\Phi(z')|s'=1], \quad E[\Phi(z')|s'=2]
        %\end{gather*}
    \item Then weight the conditional expectations by their corresponding likelihood. %The conditional expectations are:
  %  \begin{gather*}
    %  E\left[\Phi(z')|s = i\right] = p_{i1}E[\Phi(z')|s'=1] + p_{i2}E[\Phi(z')|s'=2]
   % \end{gather*}
  \end{itemize}
\end{enumerate}

@@

\begin{itemize}\setlength{\itemsep}{8pt}
  \item<1-|handout:1> Used to approximate an exogenous $AR(1)$ process
  \item <2-|handout:1>Kopecky and Suen (2010) show the Rouwenhorst method outperforms other approximations of an $AR(1)$ process
  \item <3-|handout:1>The approximation is a Markov switching process like the time-varying intercept example, but with $n$ states
  \item <4-|handout:1>The method determines the bounds of the exogenous state variables, the nodes, and the transition probabilities
  \item <5-|handout:1>Let $z\sim AR(1)$ with persistence $\rho$, mean $\mu_z$, and variance %Idea for approximation of an AR(1) process
  \begin{gather*}
    \sigma_z^2 = \sigma_\varepsilon^2/(1-\rho^2).
  \end{gather*}
  \item   <6-|handout:1>The $n$ states for the discretized process $z$ are evenly spaced on $[\mu_z-\sigma_z\sqrt{n-1},\mu_z+\sigma_z\sqrt{n-1}]$
\end{itemize}

@@

\begin{itemize}\setlength{\itemsep}{14pt}
  \item <1-|handout:1>Any calculations that are not dependent on the results of other calculations can be performed in parallel (e.g., solving for policy values at each node in the state space)
  \item <2-|handout:1>We replace \texttt{for} loops with \texttt{parfor} loops where applicable. This tells MATLAB to distribute each step in the loop across the specified number of processors
\end{itemize}

@@

%===============================================================================================================
\begin{frame}\frametitle{Conclusion}
\begin{itemize}\setlength{\itemsep}{10pt}
\item <1-|handout:1>Key takeaways:
\begin{itemize}\setlength{\itemsep}{4pt}
\item In the model without capital, regime-indexed policy functions were smoother and solution algorithm was faster
\item In model with capital,  the regime-indexed policy functions were more nonlinear and solution algorithm was slower
\end{itemize}
\item <2-|handout:1>Extensions:
\begin{itemize}\setlength{\itemsep}{4pt}
\item Solve the GHLS solution method using Smolyak discretization methods and Chebyshev polynomials 
\item This would be expected to speed up the solution algorithm, but may lead to a less accurate solution
\end{itemize}
\end{itemize}
\end{frame}
